---
title: " U.S. National Oceanic and Atmospheric Administration's (NOAA) data analize"
output: pdf_document
---


## Synopsis

TODO synopsis  

## Data Processing
Load necessary libraries.
```{r load libraries, warning=FALSE, message=FALSE}
library(ggplot2)
library(cowplot)

```
Download and read csv file.
```{r setup, cache = TRUE}
url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
file = "Data/FStormData.csv.bz2"
if(!file.exists(file)){
  download.file(url,destfile = file)
}
data <- read.csv(file)
```
Create data set "fatalities", which is a subset of original data with events with number of fatalities greater than 99% of data.
```{r,fatalites_subset}
fatalites <- subset(data, FATALITIES>0)
q_fatalites <-quantile(fatalites$FATALITIES, probs=c(.25,.50,0.95,0.99))
fatalites_99 <- subset(fatalites, FATALITIES>q_fatalites[4])
```
Create data set "injuries", which is a subset of original data with events with number of injuries greater than 99% of data.
```{r,injuriess_subset}
injuries <- subset(data, INJURIES>0)
q_injuries <-quantile(injuries$INJURIES, probs=c(.25,.50,0.95,0.99))
injuries_99 <- subset(injuries, INJURIES>q_injuries[4])
```

##Results

### Witch events are most harmful.

```{r harmfuless}

f <- ggplot(fatalites_99, aes(x=log(FATALITIES),y=EVTYPE)) + 
      geom_bar(fill="red",stat="identity" ) + 
      ggtitle("Fatalities per event type") + 
      xlab("log(Fatalities)") +
      ylab("Event type") +
      theme_dark()

i <- ggplot(injuries_99, aes(x=log(INJURIES),y=EVTYPE)) + 
      geom_bar(fill="red",stat="identity" ) + 
      ggtitle("Injuries per event type") + 
      xlab("log(Injuries)") +
      ylab("Event type") +
      theme_dark()

plot_grid(f,i,nrow = 2, ncol = 1)
```
Fatalities
```{r}
print(quantile(fatalites$FATALITIES))
```
Injuries
```{r}
print(quantile(injuries$INJURIES))
```
As we can see vast majority of fatalities and injures is caused by less than 1% of events.The most harmful ones are: tornadoes, floods, excessive and extreme heat and head in general. One interesting thing worth noting is that floods cause lots of the injures but almost non of the fatalities.  

### Witch events have the greatest economic consequences.
TODO



